---
title: 'Predictive analytics: `r params$TITLE`'
author: "Stylianos Serghiou"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
params:
  PROPORTION_OF_DATA:
    label: "Proportion of data to use"
    value: 1
    input: slider
    min: 0
    max: 1
    step: 0.05
    sep: ""
  CONFIG: 
    label: "Config file name"
    value: "config_is-poor-outcome.R"
    input: select
    choices: [
      "config_is-poor-outcome.R",
      "config_is-poor-outcome_haemTRUE.R",
      "config_is-poor-outcome_haemFALSE.R",
      "config_has-complication-major.R",
      "config_has-complication-minor.R",
      "config_is-obliterated.R"
    ]
  TITLE:
    label: "Outcome title"
    value: "Poor outcomes"
    input: select
    choices: [
      "Poor outcome",
      "Poor outcome - Hemorrhage",
      "Poor outcome - No Hemorrhage",
      "Complications - Major",
      "Complications - Minor",
      "Obliterated"
    ]
output:
  rmdformats::readthedown:
    highlight: kate
    df_print: kable    # obviates %>% kable; does not replace styling though
    code_folding: hide # [hide, show] (comment out to not give option)
  pdf_document:
    highlight: tango
    df_print: kable
    latex_engine: pdflatex
    keep_tex: yes
  tufte::tufte_handout: default
  tufte::tufte_html: 
    toc: TRUE
  prettydoc::html_pretty:
    # no cold_folding available
    theme: hpstr      # or: architect; https://github.com/yixuan/prettydoc
    highlight: github # or: vignette
    toc: TRUE         # no toc_float available
    df_print: kable   # obviates %>% kable; does not replace styling though
header-includes:
- \DeclareUnicodeCharacter{3B8}{~}
- \DeclareUnicodeCharacter{3B1}{~}
- \DeclareUnicodeCharacter{3B2}{~}
- \DeclareUnicodeCharacter{223C}{~}
- \DeclareUnicodeCharacter{2264}{~}
- \DeclareUnicodeCharacter{2265}{~}
- \DeclareUnicodeCharacter{2581}{~}
- \DeclareUnicodeCharacter{2582}{~}
- \DeclareUnicodeCharacter{2583}{~}
- \DeclareUnicodeCharacter{2585}{~}
- \DeclareUnicodeCharacter{2586}{~}
- \DeclareUnicodeCharacter{2587}{~}
- \DeclareUnicodeCharacter{FB00}{~}
- \usepackage{graphicx}
editor_options: 
  chunk_output_type: inline
---

<style>
p {

text-align: justify;
text-justify: interword;
padding: 0 0 0.5em 0

}
</style>

```{r knitr, echo=FALSE}
# Load packages
library(knitr)
library(rmdformats)
library(kableExtra)
library(ggplot2)
library(magrittr)


######### knitr

# Define chunk options
opts_chunk$set(
  echo = T,
  cache = F, # if TRUE, no need to rerun chunks
  # cache.lazy = TRUE,  # use with big objects (>1 GB)
  cache.comments = F, # do not rebuild if comments change
  tidy = F, # can play with this
  warning = F,
  message = F,
  comment = NA,
  fig.align = "center",
  fig.width = 7,
  fig.path = file.path(tempdir(), "../../outputs/predictive-analytics/figs/"),
  # fig.path = "../../outputs/predictive-analytics/figs/", # export all figures
  linewidth = 91,
  width = 75
)

# Initiatialize hook
hook_output <- knit_hooks$get("output")

# Hook to wrap output text when it exceeds 'n' using linewidth
knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$linewidth)) {
    x <- knitr:::split_lines(x)

    # wrap lines wider than 'n'
    if (any(nchar(x) > n)) {
      x <- strwrap(x, width = n)
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})

# Times a chunk and prints the time it took to run it under the chunk
# To time a chunk, include in the chunk options: {r my_chunk, timeit=TRUE}
knitr::knit_hooks$set(timeit = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste("Time for this code chunk:", as.character(res))
    }
  }
}))

# For more knitr options visit: https://yihui.name/knitr/options/
# and his github page: https://github.com/yihui/knitr-examples


######### kableExtra

options(
  knitr.kable.NA = "", # replace NAs in tables with blank
  digits = 5 # round digits (doesn't work without this here!)
)

## Example use
# data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9)) %>%
#   kable(booktabs = T) %>% kable_styling()

# Function to simplify table styling
sable <- function(tab, escape = T, full_width = F, drop = F, font_size = 12) {
  if (drop) {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      collapse_rows(valign = "top") %>%
      kable_styling(
        "striped",
        position = "center",
        full_width = full_width,
        font_size = font_size
      )
  } else {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      kable_styling(
        "striped",
        position = "center",
        full_width = full_width,
        font_size = font_size
      )
  }
}

## Guidelines
# No longer need to define options(knitr.table.format = "html"). It is now automatically done as soon as you load kableExtra
# No need to run kable() every time - done automatically as soon as you load kableExtra
# Loading kableExtra nullifies any styling applied by df_table: kable in the preamble - if you are content with standard formatting, DO NOT load kableExtra


#########  ggplot2

# Set up preferred theme in ggplot2
my_theme <-
  # this and theme_minimal() are my favorite
  theme_light() +
  theme(
    axis.ticks = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y = element_text(margin = margin(r = 7)),
    legend.key = element_rect(colour = NA, fill = NA), # Avoid borders
    panel.border = element_blank(),
    text = element_text(color = "grey20"),
    title = element_text(face = "bold")
  )

# Make the above theme the default theme
original_theme <- theme_set(my_theme)

# Use ggplot2::ggsave to save plots after plotting - this reduces size dramatically


######### Tabbed sections

# You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

## Quarterly Results {.tabset}

### By Product


######### Update package

# To update the package use:
# devtools::install_github(serghiou/serghioutemplates)
```


# Thoughts

## Questions

- What deficit are we talking about? Is it covered by the mRS?
- Should the pre-treatment mRS be in the model? Already covered by the others?


## Backlog

- Honest estimation
- G estimation (using only significant ones to get accurate ORs)
- Make sure that the number of outcomes (26) is valid.
- Bootstrap the whole process - how many models include each of the features?
- The SuSiE method: https://www.biorxiv.org/content/10.1101/501114v1
- Choose best model with forward/backward exclusion
- Feature selection with honest estimation + p-value-based + bootstrap
- Feature selection with honest estimation + LASSO-based + bootstrap
- Implement PCA to create independent variables and not drop variables
- Bayesian fitting
- Change to use tidybayes and tidymodels
- Use neural networks to fit
- Consider a time-to-event analysis


------------------------------------------------------------------------

# Setup

## Imports

```{r setup}
# Load packages
library(furrr)
library(glmnet)
library(magrittr)
library(patchwork)
library(selectiveInference)
library(tidyverse)

# Source code

```


## Parameters

```{r}
# Source outcome-specific configs
source(params$CONFIG)

# Destination locations
DST_DIRNAME <- paste0("../../outputs/")

# Source locations
SRC_DIRNAME <- "../../data/3_tidy/patients"
SRC_BASENAME <- "patients-daily_2024-07-21.csv"

# Should the Dockerfile be automatically updated?
UPDATE_DOCKERFILE <- FALSE

# Parallelize - off by default because it's implemented in the renderer
PARALLELIZE <- FALSE

# Set the seed
set.seed(1891)
```


## Variables

```{r}
EXPOSURES_CONTINUOUS <- c(
  "Age at 1st treatment (years)" = "age_at_first_treatment_yrs",
  "mRS (presentation)" = "modified_rankin_score_presentation",
  "mRS (pre-treatment)" = "modified_rankin_score_pretreatment",
  "mRS (1-week post-op)" = "modified_rankin_score_postop_within_1_week",
  "mRS (final)" = "modified_rankin_score_final",
  "Nidus size (cm)" = "max_size_cm",
  "Spetzler-Martin grade" = "spetzler_martin_grade",
  "Lawton-Young grade" = "lawton_young_grade",
  "Size score" = "size_score"
)

EXPOSURES_BINARY <- c(
  "Sex (Male)" = "is_male",
  "First tx after Jun 2009" = "is_first_tx_after_2009",
  "First tx as of 2001 onwards" = "is_first_tx_from_2001",
  "Involves eloquent location" = "is_eloquent_location",
  "Has deep venous drainage" = "has_deep_venous_drainage",
  "Diffuse nidus" = "is_diffuse_nidus",
  "Hemorrhage at presentation" = "has_hemorrhage",
  "Seizures at presentation" = "has_seizures",
  "Deficit at presentation" = "has_deficit",
  "Paresis at presentation" = "has_paresis",
  "Presence of aneurysms" = "has_associated_aneurysm",
  "Spetzler-Martin grade < 4" = "is_spetzler_martin_grade_less_than_4",
  "Had surgery" = "is_surgery"
)

EXPOSURES_CATEGORICAL <- c(
  "Location" = "location",
  "Venous drainage" = "venous_drainage",
  "Modality of treatment" = "procedure_combinations"
)

OUTCOMES <- c(
  "Poor outcome (mRS >= 3)" = "is_poor_outcome",
  "Obliteration" = "is_obliterated",
  "Complications - minor" = "has_complication_minor",
  "Complications - major" = "has_complication_major",
  "mRS change (final - pre-treatment)" =
    "modified_rankin_score_final_minus_pretx",
  "mRS change (final - presentation)" =
    "modified_rankin_score_final_minus_presentation",
  "mRS change direction (Final - Pre-tx)" = "change_in_mrs_tx_vs_final"
)
```


## Outputs

Create lists.

```{r}
plots <- list()
tables <- list()
models <- list()
```


## Functions

### R utils

```{r}
```


### Data analysis utils

Parse inconsistent dates.

```{r}
# Function to parse dates using the appropriate format
# (needed b/c some are "%m/%d/%Y" and some "%m/%d/%y")
parse_inconsistent_dates <- function(date_col) {
  # Try %m/%d/%y format first
  parsed_dates <- as.Date(date_col, format = "%m/%d/%Y")
  
  # Identify and correct dates before 1950
  parsed_dates <- ifelse(
    !is.na(parsed_dates) & parsed_dates < as.Date("1950-01-01"),
    as.Date(date_col, format = "%m/%d/%y"),
    parsed_dates
  )
  
  # Return corrected dates
  as.Date(parsed_dates, origin = "1970-01-01")
}
```


Descriptive statistics - continuous variables.

```{r}
compute_descriptive_continuous <- function(
    df = df_uni,
    cols = c(EXPOSURES_CONTINUOUS),
    outcome_col = CHOSEN_OUTCOME,
    outcome_levels = CHOSEN_OUTCOME_LEVELS,
    outcome_labels = CHOSEN_OUTCOME_LABELS,
    use_interaction = TRUE,
    interaction_col = "has_hemorrhage",
    interaction_col_name = "Hemorrhage at presentation",
    interaction_col_value = TRUE,
    prefix = "Haem") {
  
  # Apply desired formatting to numbers
  format_numbers <- function(x) {
    sprintf("%.1f", x)
  }

  # Apply interaction
  # (keep this selection here to avoid impact of prev. defined interaction_col)
  if (use_interaction) {
    # Compute new dataset
    df <-
      df %>%
      select(cols, outcome_col, interaction_col) %>%
      drop_na(outcome_col, interaction_col) %>% 
      filter(!!sym(interaction_col) == interaction_col_value)
  } else {
    # Compute new dataset
    df <-
      df %>%
      select(cols, outcome_col) %>%
      drop_na(outcome_col)
  }

  # Compute descriptive statistics
  desc_stats <-
    df %>%
    pivot_longer(where(is.numeric), names_to = "keys", values_to = "values") %>%
    group_by(!!sym(outcome_col), keys) %>%
    summarize(
      num_total = n(),
      num_missing = sum(is.na(values)),
      mean = mean(values, na.rm = TRUE),
      sd = sd(values, na.rm = TRUE),
      min = quantile(values, 0, na.rm = TRUE),
      max = quantile(values, 1, na.rm = TRUE),
      q_50 = quantile(values, 0.50, na.rm = TRUE),
      q_25 = quantile(values, 0.25, na.rm = TRUE),
      q_75 = quantile(values, 0.75, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(
      stats = glue::glue("{format_numbers(mean)} ({format_numbers(sd)})"),
      !!sym(outcome_col) := factor(
        !!sym(outcome_col),
        levels = outcome_levels,
        labels = outcome_labels
      )
    ) %>%
    pivot_wider(
      id_cols = keys,
      names_from = !!sym(outcome_col),
      values_from = stats
    ) %>%
    relocate(outcome_labels[1], .after = outcome_labels[2])

  # Compute p-values
  pvals <-
    df %>%
    pivot_longer(where(is.numeric), names_to = "keys", values_to = "values") %>%
    group_by(keys) %>%
    summarize(
      pvalue = wilcox.test(values ~ !!sym(outcome_col))$p.value
    )

  # Synthesize
  out <-
    desc_stats %>%
    left_join(pvals)

  # Add sample size to column names
  # Count each outcome group by its OWN label (robust to level/label order)
  n_by_label <- setNames(
    c(
      sum(df %>% pull(outcome_col) == outcome_levels[1], na.rm = TRUE),
      sum(df %>% pull(outcome_col) == outcome_levels[2], na.rm = TRUE)
    ),
    outcome_labels
  )
  # Build headers so each label carries its own n (order matches relocate())
  new_col_names <- c(
    glue::glue("{prefix} - {outcome_labels[2]} (n={n_by_label[[outcome_labels[2]]]})"),
    glue::glue("{prefix} - {outcome_labels[1]} (n={n_by_label[[outcome_labels[1]]]})"),
    glue::glue("{prefix} - P-value")
  )

  out <-
    out %>%
    rename_with(
      .cols = !!sym(outcome_labels[2]):pvalue, ~new_col_names
    ) %>%
    mutate(
      keys = ifelse(keys == interaction_col, interaction_col_name, keys)
    )

  return(out)
}
```


Descriptive statistics - binary variables.

```{r}
compute_descriptive_binary <- function(
    df = df_uni,
    cols = c(EXPOSURES_BINARY),
    outcome_col = CHOSEN_OUTCOME,
    outcome_levels = CHOSEN_OUTCOME_LEVELS,
    outcome_labels = CHOSEN_OUTCOME_LABELS,
    use_interaction = TRUE,
    interaction_col = "has_hemorrhage",
    interaction_col_name = "Hemorrhage at presentation",
    interaction_col_value = TRUE,
    prefix = "Haem") {
  
  # Define arguments
  cols <- c(cols[!cols == interaction_col])

  # Apply desired formatting to numbers
  format_numbers <- function(x, decimals = 0) {
    decimals <- paste0("%.", decimals, "f")
    sprintf(decimals, x)
  }

  # Compute new dataset
  df <-
    df %>%
    select(all_of(cols), outcome_col, interaction_col) %>%
    drop_na(outcome_col, interaction_col)

  # Apply interaction
  if (use_interaction) {
    df <-
      df %>%
      filter(!!sym(interaction_col) == interaction_col_value)
  }

  # Compute descriptive statistics
  desc_stats <-
    df %>%
    pivot_longer(
      cols = -c(outcome_col),
      names_to = "keys",
      values_to = "values"
    ) %>%
    group_by(!!sym(outcome_col), keys) %>%
    summarize(
      num_total = n(),
      num_missing = sum(is.na(values)),
      num_with = sum(values, na.rm = TRUE),
      num_without = sum(!values, na.rm = TRUE),
      pct_with = mean(values, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(
      stats = glue::glue("{num_with} ({format_numbers(pct_with*100)}%)"),
      !!sym(outcome_col) := factor(
        !!sym(outcome_col),
        levels = outcome_levels,
        labels = outcome_labels
      )
    ) %>%
    pivot_wider(
      id_cols = keys,
      names_from = !!sym(outcome_col),
      values_from = stats
    ) %>%
    relocate(outcome_labels[1], .after = outcome_labels[2])

  # Compute p-values
  pvals <-
    df %>%
    pivot_longer(
      cols = -c(outcome_col),
      names_to = "keys",
      values_to = "values"
    ) %>%
    group_by(keys) %>%
    summarize(
      pvalue = kruskal.test(values ~ !!sym(outcome_col))$p.value
    )

  # Synthesize
  out <-
    desc_stats %>%
    left_join(pvals)

  # Add sample size to column names
  # Count each outcome group by its OWN label (robust to level/label order)
  n_by_label <- setNames(
    c(
      sum(df %>% pull(outcome_col) == outcome_levels[1], na.rm = TRUE),
      sum(df %>% pull(outcome_col) == outcome_levels[2], na.rm = TRUE)
    ),
    outcome_labels
  )
  # Build headers so each label carries its own n (order matches relocate())
  new_col_names <- c(
    glue::glue("{prefix} - {outcome_labels[2]} (n={n_by_label[[outcome_labels[2]]]})"),
    glue::glue("{prefix} - {outcome_labels[1]} (n={n_by_label[[outcome_labels[1]]]})"),
    glue::glue("{prefix} - P-value")
  )

  out <-
    out %>%
    rename_with(
      .cols = !!sym(outcome_labels[2]):pvalue, ~new_col_names
    ) %>%
    mutate(
      keys = ifelse(keys == interaction_col, interaction_col_name, keys)
    )

  return(out)
}
```

### Predictive analytics utils




------------------------------------------------------------------------

# Read

```{r}
# File path
filepath <- file.path(SRC_DIRNAME, SRC_BASENAME)

# Read
patients_ <-
  read_csv(
    filepath,
    col_types = cols(
      .default = col_guess(), # Default behavior for other columns
      date_first_treatment = col_character()
    )
  ) %>%
  dplyr::sample_frac(params$PROPORTION_OF_DATA)

# If analyzing a subset of the data, restrict the dataset
if (params$TITLE == "Poor outcome - Hemorrhage") {
  patients_ <- patients_ %>% filter(has_hemorrhage)
}
if (params$TITLE == "Poor outcome - No Hemorrhage") {
  patients_ <- patients_ %>% filter(!has_hemorrhage)
}
if (params$TITLE == "Poor outcome - High grade") {
  patients_ <- patients_ %>% filter(spetzler_martin_grade >= 4)
}
if (params$TITLE == "Poor outcome - Low grade") {
  patients_ <- patients_ %>% filter(spetzler_martin_grade < 4)
}
```


------------------------------------------------------------------------

# Conform

## Setup

Remove all empty rows.

```{r}
patients_ <-
  patients_ %>%
  filter(!is.na(patient_id)) %>%
  arrange(patient_id)
```


Create two dataframes - one for univariable and one for multivariable analysis. This is because variables for multivariable analysis will be recoded to reduce the number of levels to avoid overfitting the model.

```{r}
df_uni <-
  patients_ %>%
  filter(is_eligible) %>% 
  filter(!is.na(spetzler_martin_grade))

df_multi <-
  patients_ %>%
  filter(is_eligible) %>% 
  filter(!is.na(spetzler_martin_grade))
```


## Recode columns

For `venous_drainage` if "Both", mark as "Deep."

```{r}
df_multi <-
  df_multi %>%
  mutate(
    venous_drainage = ifelse(venous_drainage == "Both", "Deep", venous_drainage)
  )
```


For `procedure_combinations`, change > 1 to multimodal to reduce levels.

```{r}
df_multi <-
  df_multi %>%
  mutate(procedure_combinations = case_when(
    nchar(procedure_combinations) > 1 ~ "Multimodal",
    .default = procedure_combinations
  ))
```


For `procedure_combinations`, indicate if surgery-based or not.

```{r}
df_multi <-
  df_multi %>%
  mutate(
    is_surgery = str_detect(procedure_combinations, "S")
  ) %>%
  relocate(is_surgery, .after = procedure_combinations)

df_uni <-
  df_uni %>%
  mutate(
    is_surgery = str_detect(procedure_combinations, "S")
  ) %>%
  relocate(is_surgery, .after = procedure_combinations)
```


For `location`, reduce levels (use "Other").

```{r}
df_multi <-
  df_multi %>%
  mutate(
    location = ifelse(location == "Corpus Callosum", "Other", location),
    location = ifelse(location == "Cerebellum", "Other", location),
    # location = ifelse(location == "Deep", "Other", location),
    location = factor(location),
    location = relevel(location, ref = "Other")
  )
```


For `spetzler_martin_grade`, create a binary variant of 1-3 vs. 4-5.

```{r}
# For multivariable analysis
df_multi <-
  df_multi %>%
  mutate(
    # Divides into low grade vs. high grade
    is_spetzler_martin_grade_less_than_4 = spetzler_martin_grade < 4
  ) %>%
  relocate(is_spetzler_martin_grade_less_than_4, .after = spetzler_martin_grade)

# For univariable analysis
df_uni <-
  df_uni %>%
  mutate(
    is_spetzler_martin_grade_less_than_4 = spetzler_martin_grade < 4
  ) %>%
  relocate(is_spetzler_martin_grade_less_than_4, .after = spetzler_martin_grade)
```


For `date_first_treatment`,transform into dates and create > 2009 vs. not.

```{r}
# Parse into Date format and create dichotomous variable
df_uni <-
  df_uni %>%
  mutate(across(starts_with("date"), ~ parse_inconsistent_dates(.x))) %>% 
  mutate(is_first_tx_after_2009 = 
           date_first_treatment >= as.Date("2009-06-30")) %>%
  mutate(is_first_tx_from_2001 =
           date_first_treatment >= as.Date("2001-01-01"))

# Fix dates and parse into Date format
df_multi <-
  df_multi %>%
  mutate(across(starts_with("date"), ~ parse_inconsistent_dates(.x))) %>% 
  mutate(is_first_tx_after_2009 = 
           date_first_treatment >= as.Date("2009-06-30")) %>%
  mutate(is_first_tx_from_2001 =
           date_first_treatment >= as.Date("2001-01-01"))

# Check for missingness
num_missing <-
  df_uni %>% 
  filter(is.na(date_first_treatment)) %>% 
  pull(patient_id) %>% 
  length()

ids_missing <- 
  df_uni %>%
  filter(is.na(date_first_treatment)) %>% 
  pull(patient_id)

# Check for non-sensical data
min(df_uni$date_first_treatment, na.rm = T) > as.Date("1980-01-01")
max(df_uni$date_first_treatment, na.rm = T) < as.Date("2025-01-01")
```


## Missingness

```{r}
# Get cols
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  "is_first_tx_after_2009",
  "is_first_tx_from_2001",
  EXPOSURES_CATEGORICAL,
  OUTCOMES
))

# Count missingness
df_multi %>%
  select(cols) %>%
  summarise_all(~ sum(is.na(.))) %>%
  pivot_longer(everything(), values_to = "num_missing") %>%
  arrange(desc(num_missing)) %>%
  filter(num_missing > 0) %>%
  sable()
```


Which eligible patients are missing each variable?

```{r}
df_multi %>%
  filter(is_eligible) %>%
  select(mrn, cols) %>%
  mutate(across(-mrn, is.na)) %>%
  pivot_longer(
    cols = -mrn,
    names_to = "name",
    values_to = "is_missing"
  ) %>%
  filter(is_missing) %>%
  group_by(name) %>%
  summarize(mrn = str_c(mrn, collapse = ", ")) %>%
  sable()
```


------------------------------------------------------------------------

# Visualizations

## Overall

Distribution of values of the exposures across levels of the outcome.

```{r, fig.width=12, fig.height=8}
# Define the predictors of interest
cols <- c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  CHOSEN_OUTCOME
)

# Reshape the data
df_long <-
  df_uni %>%
  select(all_of(cols)) %>%
  pivot_longer(-CHOSEN_OUTCOME, names_to = "predictor", values_to = "value")

# Plot the box plots
df_long %>%
  ggplot(aes_string(
    x = CHOSEN_OUTCOME,
    y = "value",
    fill = CHOSEN_OUTCOME,
    color = CHOSEN_OUTCOME
  )) +
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~predictor, scales = "free") +
  labs(x = "Outcome", y = "Value", fill = "Outcome") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 11),
    legend.position = "none"
  )
```


## By hemorrhage

```{r, fig.width=14, fig.height=12}
# Define the predictors of interest
cols <- c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  CHOSEN_OUTCOME
)

# Reshape the data
df_long <-
  df_uni %>%
  select(all_of(cols)) %>%
  pivot_longer(
    cols = -c(CHOSEN_OUTCOME, `Hemorrhage at presentation`),
    names_to = "predictor",
    values_to = "value"
  )

# Plot the box plots
df_long %>%
  ggplot(aes_string(
    x = CHOSEN_OUTCOME,
    y = "value",
    fill = "`Hemorrhage at presentation`",
    color = "`Hemorrhage at presentation`"
  )) +
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~predictor, scales = "free") +
  labs(x = "Outcome", y = "Value", fill = "Outcome") +
  theme_minimal() +
  guides(fill = "none") +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 11)
  )
```


## By SM grade high vs. low

```{r, fig.width=14, fig.height=12}
# Define the predictors of interest
cols <- c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  CHOSEN_OUTCOME
)

# Reshape the data
df_long <-
  df_uni %>%
  select(all_of(cols)) %>%
  pivot_longer(
    cols = -c(CHOSEN_OUTCOME, `Spetzler-Martin grade < 4`),
    names_to = "predictor",
    values_to = "value"
  )

# Plot the box plots
df_long %>%
  ggplot(aes_string(
    x = CHOSEN_OUTCOME,
    y = "value",
    fill = "`Spetzler-Martin grade < 4`",
    color = "`Spetzler-Martin grade < 4`"
  )) +
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~predictor, scales = "free") +
  labs(x = "Outcome", y = "Value", fill = "Outcome") +
  theme_minimal() +
  guides(fill = "none") +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 11)
  )
```


## Specialty

### SM grade vs. complications

```{r}
# Rename
complications <- c(
  has_complication_minor = "Minor",
  has_complication_major = "Major"
)

# Create table for plotting
t <- 
  df_uni %>% 
  filter(is_eligible) %>% 
  select(
    patient_id,
    spetzler_martin_grade,
    has_complication_minor,
    has_complication_major
  ) %>% 
  pivot_longer(starts_with("has_complication")) %>% 
  group_by(spetzler_martin_grade, name) %>% 
  summarize(
    num_patients = sum(value, na.rm = T)
  ) %>% 
  group_by(name) %>% 
  mutate(pct_patients = num_patients / sum(num_patients))

# Create figure
plots$sm_grade_vs_complications_bar_chart <- 
  t %>%
  mutate(name = factor(name, names(complications), complications)) %>% 
  ggplot(aes(
    x = spetzler_martin_grade,
    y = pct_patients,
    color = name,
    fill = name)
  ) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7
  ) +
  geom_label(
    aes(label = num_patients),
    position = position_dodge(width = 0.8),
    vjust = -0.2,
    size = 3,
    fill = "white",
    label.size = NA,  # Removes the label border
    show.legend = FALSE
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_fill_manual(values = c("Minor" = "#00BFC4", "Major" = "#F8766D")) +
  scale_color_manual(values = c("Minor" = "#00BFC4", "Major" = "#F8766D")) +
  labs(
    x = "Spetzler-Martin Grade",
    y = "Patients (%)",
    fill = "Complications",
    color = "Complications"
  ) +
  theme(
    legend.position = c(0.3, 0.95),
    legend.justification = c("right", "top")
  )
  
# Plot
plots$sm_grade_vs_complications_bar_chart
```


## By date

Every 1 year.

```{r}
# Create table
df_by_date <-
  df_uni %>%
  filter(!is.na(date_first_treatment)) %>% 
  mutate(year_first_treatment = format(date_first_treatment, "%Y")) %>% 
  group_by(year_first_treatment) %>% 
  summarize(
    num_patients = n(),
    num_poor_outcome = sum(is_poor_outcome, na.rm = T),
    pct_poor_outcome = 100 * num_poor_outcome / num_patients
  )

df_by_date %>% 
  ggplot(aes(x = year_first_treatment, y = pct_poor_outcome)) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(num_poor_outcome, "/", num_patients)
    ),
    size = 2,
    vjust = -0.5  # Adjust vertical position above the bar
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1.5, vjust = 2.5)
  ) +
  labs(
    x = "Year of First Treatment",
    y = "Poor outcomes (%)",
    title = "Percentage of Poor Outcomes by Year"
  )
```


Every 5 years.

```{r}
# Create table
df_by_date <-
  df_uni %>%
  filter(!is.na(date_first_treatment)) %>% 
  mutate(year_first_treatment = as.numeric(format(date_first_treatment, "%Y"))) %>%
  mutate(
    year_interval = paste0(
      floor(year_first_treatment / 5) * 5,
      "-",
      floor(year_first_treatment / 5) * 5 + 4
    )
  ) %>% 
  group_by(year_interval) %>% 
  summarize(
    num_patients = n(),
    num_poor_outcome = sum(is_poor_outcome, na.rm = T),
    pct_poor_outcome = 100 * num_poor_outcome / num_patients
  )

# Visualize
df_by_date %>% 
  ggplot(aes(x = year_interval, y = pct_poor_outcome)) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(num_poor_outcome, "/", num_patients)
    ),
    size = 3,
    vjust = -0.5  # Adjust vertical position above the bar
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 1)
  ) +
  labs(
    x = "Year of First Treatment",
    y = "Poor outcomes (%)",
    title = "Percentage of Poor Outcomes by Year"
  )
```


Every 5 years.

```{r, fig.height=6, fig.width=6}
# Create table
df_by_date <-
  df_uni %>%
  filter(!is.na(date_first_treatment)) %>% 
  mutate(year_first_treatment = as.numeric(format(date_first_treatment, "%Y"))) %>%
  mutate(
    year_interval = paste0(
      floor(year_first_treatment / 10) * 10,
      "-",
      floor(year_first_treatment / 10) * 10 + 9
    )
  ) %>% 
  group_by(year_interval) %>% 
  summarize(
    num_patients = n(),
    num_poor_outcome = sum(is_poor_outcome, na.rm = T),
    pct_poor_outcome = 100 * num_poor_outcome / num_patients
  )

# Visualize
df_by_date %>% 
  ggplot(aes(x = year_interval, y = pct_poor_outcome)) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(num_poor_outcome, "/", num_patients)
    ),
    size = 3,
    vjust = -0.5  # Adjust vertical position above the bar
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 1)
  ) +
  labs(
    x = "Year of First Treatment",
    y = "Poor outcomes (%)",
    title = "Percentage of Poor Outcomes by Year"
  )
```


------------------------------------------------------------------------

# Descriptive statistics

## Setup

```{r}

```


## Non-specific

```{r}
# Rename dataframe
`Pediatric AVMs` <-
  df_uni %>%
  filter(is_eligible) %>%
  select(-is_eligible, -comments)

# Create summary
`Pediatric AVMs` %>%
  summarytools::dfSummary(display.labels = FALSE) %>%
  print(
    file =
      "../../outputs/descriptive-statistics/descriptive_statistics_eligible.html",
    footnote = NA
  )

# Remove unwanted dataframe
remove(`Pediatric AVMs`)
```


## By hemorrhage - table1

https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html

```{r eval=FALSE}
library(table1)

table1::table1(
  ~ factor(is_male) + age_at_first_treatment_yrs + factor(is_eloquent_location) + max_size_cm | has_hemorrhage,
  data = df_uni
)
```

```{r eval=FALSE}
# Initialize values
df <- df_uni

# Remove missing values in stratification variables
df <-
  df %>%
  drop_na(has_hemorrhage, CHOSEN_OUTCOME)

# Define columns
cols <- c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  OUTCOMES[OUTCOMES == CHOSEN_OUTCOME]
)

# Assign labels to the variables in the dataframe
for (var in cols) {
  label(df[[var]]) <- names(cols[cols == var])
}

# Assign meaningful labels to the has_hemorrhage variable
df$has_hemorrhage <- factor(
  df$has_hemorrhage,
  levels = c(FALSE, TRUE),
  labels = c("No Hemorrhage", "Has Hemorrhage")
)

df[, CHOSEN_OUTCOME] <- factor(
  df %>% pull(CHOSEN_OUTCOME),
  levels = CHOSEN_OUTCOME_LEVELS,
  labels = CHOSEN_OUTCOME_LABELS
)

# Define custom rendering functions if needed
render_continuous <- function(x) {
  with(stats.apply.rounding(stats.default(x), digits = 2), c(
    "Mean (SD)" = sprintf("%s (&plusmn; %s)", MEAN, SD),
    "Median (IQR)" = sprintf("%s (%s - %s)", MEDIAN, Q1, Q3)
  ))
}

compute_pvalues <- function(x, ...) {
  # Construct vectors of data y, and groups (strata) g
  y <- unlist(x)
  g <- factor(rep(1:length(x), times = sapply(x, length)))
  if (is.numeric(y)) {
    # For numeric variables, perform a standard 2-sample t-test
    p <- t.test(y ~ g)$p.value
  } else {
    # For categorical variables, perform a chi-squared test of independence
    p <- chisq.test(table(y, g))$p.value
  }
  # Format the p-value, using an HTML entity for the less-than sign.
  # The initial empty string places the output on the line below the variable label.
  c("", sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001)))
}

# Create the descriptive table
frla <- as.formula(paste("~ . | has_hemorrhage *", CHOSEN_OUTCOME))

# Create table
table1_descriptive_statistics <-
  table1(frla,
    data = df[, unname(cols)],
    render.continuous = render_continuous,
    overall = c(left = "Overall"),
    topclass = "Rtable1-zebra"
  )

# Print
table1_descriptive_statistics
```


## By hemorrhage

```{r}
if (!str_detect(params$TITLE, "Hemorrhage")) {
  
  # Define arguments
  interaction_col <- "has_hemorrhage"
  interaction_col_name <- "Hemorrhage at presentation"
  prefix_true <- "Haem"
  prefix_false <- "No Haem"
  
  # Get all
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = F,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = "All"
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = F,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = "All"
  )
  out_all <- bind_rows(out)
  
  # Get with hemorrhage
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = prefix_true
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = prefix_true
  )
  out_haem <- bind_rows(out)
  
  # Get without hemorrhage
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = F,
    prefix = prefix_false
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = F,
    prefix = prefix_false
  )
  out_no_haem <- bind_rows(out)
  
  # Synthesize
  out_complete <-
    out_all %>%
    left_join(out_haem, by = "keys") %>%
    left_join(out_no_haem, by = "keys") %>%
    arrange(`All - P-value`)
  
  # Print
  out_complete %>%
    sable()
  
}
```


## By diffuse nidus

```{r}
# Define arguments
interaction_col <- "is_diffuse_nidus"
interaction_col_name <- "Is diffuse nidus"
prefix_true <- "Diffuse"
prefix_false <- "Not Diffuse"

# Get all
out <- list()
out[[1]] <- compute_descriptive_continuous(
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out[[2]] <- compute_descriptive_binary(
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out_all <- bind_rows(out)

# Get with hemorrhage
out <- list()
out[[1]] <- compute_descriptive_continuous(
  use_interaction = T,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = prefix_true
)
out[[2]] <- compute_descriptive_binary(
  use_interaction = T,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = prefix_true
)
out_haem <- bind_rows(out)

# Get without hemorrhage
out <- list()
out[[1]] <- compute_descriptive_continuous(
  use_interaction = T,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = F,
  prefix = prefix_false
)
out[[2]] <- compute_descriptive_binary(
  use_interaction = T,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = F,
  prefix = prefix_false
)
out_no_haem <- bind_rows(out)

# Synthesize
out_complete <-
  out_all %>%
  left_join(out_haem, by = "keys") %>%
  left_join(out_no_haem, by = "keys") %>%
  arrange(`All - P-value`)

# Print
out_complete %>%
  sable()
```


## By SM grade high vs. low

```{r}
if (!str_detect(params$TITLE, "grade")) {
  # Define arguments
  interaction_col <- "is_spetzler_martin_grade_less_than_4"
  interaction_col_name <- "Spetzler-Martin grade < 4"
  prefix_true <- "Low grade"
  prefix_false <- "High grade"
  
  # Get all
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = F,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = "All"
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = F,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = "All"
  )
  out_all <- bind_rows(out)
  
  # Get with
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = prefix_true
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = T,
    prefix = prefix_true
  )
  out_haem <- bind_rows(out)
  
  # Get without
  out <- list()
  out[[1]] <- compute_descriptive_continuous(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = F,
    prefix = prefix_false
  )
  out[[2]] <- compute_descriptive_binary(
    use_interaction = T,
    interaction_col = interaction_col,
    interaction_col_name = interaction_col_name,
    interaction_col_value = F,
    prefix = prefix_false
  )
  out_no_haem <- bind_rows(out)
  
  # Synthesize
  out_complete <-
    out_all %>%
    left_join(out_haem, by = "keys") %>%
    left_join(out_no_haem, by = "keys") %>%
    arrange(`All - P-value`)
  
  # Print
  out_complete %>%
    sable()
}
```


## By date

These values represent either the mean (continuous) or number and proportion (binary) of patients with the specific exposure before or after 2009. For example, for Spetzler-Martin grade < 4, the numbers represent the number and proportion of patients with SM 1-3 before vs. after 2009.

```{r}
# Define arguments
outcome_col <- "is_first_tx_after_2009"
outcome_levels <- c(TRUE, FALSE)
outcome_labels <- c("After Jun '09", "At or before Jun '09")
avoid_columns <- c("is_first_tx_from_2001", "is_first_tx_after_2009")

# Get all
out <- list()
out[[1]] <- compute_descriptive_continuous(
  outcome_col = outcome_col,
  outcome_levels = outcome_levels,
  outcome_labels = outcome_labels,
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out[[2]] <- compute_descriptive_binary(
  cols = EXPOSURES_BINARY[!EXPOSURES_BINARY %in% avoid_columns],
  outcome_col = outcome_col,
  outcome_levels = outcome_levels,
  outcome_labels = outcome_labels,
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out_all <- bind_rows(out)

# Synthesize
out_complete <-
  out_all %>%
  arrange(`All - P-value`)

# Print
out_complete %>%
  sable()
```


New analysis (as of Jul 8, 2026) to understand poor outcomes before and after 2001 by SM grade.

These values represent either the mean (continuous) or number and proportion (binary) of patients with the specific exposure before or after 2001. For example, for Spetzler-Martin grade < 4, the numbers represent the number and proportion of patients with SM 1-3 before vs. after 2001.

```{r}
# Define arguments
outcome_col <- "is_first_tx_from_2001"
outcome_levels <- c(TRUE, FALSE)
outcome_labels <- c("2001 onwards", "Before 2001")
avoid_columns <- c("is_first_tx_from_2001", "is_first_tx_after_2009")

# Get all
out <- list()
out[[1]] <- compute_descriptive_continuous(
  outcome_col = outcome_col,
  outcome_levels = outcome_levels,
  outcome_labels = outcome_labels,
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out[[2]] <- compute_descriptive_binary(
  cols = EXPOSURES_BINARY[!EXPOSURES_BINARY %in% avoid_columns],
  outcome_col = outcome_col,
  outcome_levels = outcome_levels,
  outcome_labels = outcome_labels,
  use_interaction = F,
  interaction_col = interaction_col,
  interaction_col_name = interaction_col_name,
  interaction_col_value = T,
  prefix = "All"
)
out_all <- bind_rows(out)

# Synthesize
out_complete <-
  out_all %>%
  arrange(`All - P-value`)

# Print
out_complete %>%
  sable()
```


```{r}
if (!str_detect(params$TITLE, "grade")) {
  
# One table row for a given data subset; `outcome` is a column name as a string
era_row <- function(data, group_label, outcome = CHOSEN_OUTCOME) {
  
  # Total patients excluded for any missing (era or outcome)
  missing <- sum(is.na(data$is_first_tx_from_2001) | is.na(data[[outcome]]))

  # Only include cells with data
  d <-
    data %>%
    filter(!is.na(is_first_tx_from_2001))
  
  # Fisher's exact test: era (2x2) vs outcome, within this group
  p <- fisher.test(table(d$is_first_tx_from_2001, d[[outcome]]))$p.value
  
  # Create row
  d %>%
    mutate(era = factor(
      if_else(is_first_tx_from_2001, "2001–2024", "1985–2000"),
      levels = c("1985–2000", "2001–2024")
    )) %>%
    group_by(era, .drop = FALSE) %>%
    summarise(
      n = sum(.data[[outcome]], na.rm = TRUE),
      N = sum(!is.na(.data[[outcome]])),
      .groups = "drop"
    ) %>%
    rowwise() %>%
    mutate(
      `n (%)`  = if (N > 0) sprintf("%d/%d (%.1f%%)", n, N, 100 * n / N) else "0/0",
      `95% CI` = if (N > 0) {
        ci <- binom.test(n, N)$conf.int * 100
        sprintf("%.1f–%.1f%%", ci[1], ci[2])
      } else NA_character_
    ) %>%
    ungroup() %>%
    select(era, `n (%)`, `95% CI`) %>%
    pivot_wider(
      names_from  = era,
      values_from = c(`n (%)`, `95% CI`),
      names_glue  = "{era} {.value}",
      names_vary  = "slowest"
    ) %>%
    mutate(Group = group_label, "Missing era" = missing, .before = 1) %>%
    mutate(`Fisher p` = format.pval(p, digits = 2, eps = 0.001))
}

# Create table
tbl <- bind_rows(
  era_row(df_uni, "Overall"),
  era_row(filter(df_uni, is_spetzler_martin_grade_less_than_4), "SM 1–3"),
  era_row(filter(df_uni, !is_spetzler_martin_grade_less_than_4), "SM 4–5")
)

# Print table
sable(tbl)
}
```

Interpretation: Poor functional outcome (modified Rankin Scale ≥3) was compared between treatment eras, defined by year of first treatment (1985–2000 vs 2001–2024). Overall, the rate of poor outcome was comparable across eras (6/51 [11.8%] vs 19/170 [11.2%]; Fisher's exact p = 1.00). This pattern was consistent within Spetzler–Martin (SM) grade strata. Among low-grade AVMs (SM 1–3), poor outcome occurred in 3/28 (10.7%) in the earlier era and 7/125 (5.6%) in the later era (p = 0.39). Among high-grade AVMs (SM 4–5), poor outcome occurred in 3/23 (13.0%) versus 12/45 (26.7%); although numerically higher in the later era, this difference did not reach statistical significance (p = 0.23) and was estimated with wide confidence intervals (95% CI 2.8–33.6% and 14.6–41.9%, respectively), reflecting the small number of events. No statistically significant difference in poor-outcome rate by treatment era was observed in any subgroup.


------------------------------------------------------------------------

# Associational statistics

## Correlations

Correlation matrix.

```{r, fig.height=8, fig.width=8}
# Cols
cols <- c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL,
  OUTCOMES
)

# Remove interaction terms
if (str_detect(params$TITLE, "Hemorrhage")) {
  cols <- cols[!cols == "has_hemorrhage"]
}
if (str_detect(params$TITLE, "grade")) {
  x <- c("spetzler_martin_grade", "is_spetzler_martin_grade_less_than_4")
  cols <- cols[!cols %in% x]
}
  
# Create new dataframe
df_ <-
  df_uni %>%
  select(all_of(cols))

# Convert the outcome variable to numeric for correlation calculation
df_ <-
  df_ %>%
  select(where(is.numeric), where(is.logical)) %>%
  mutate(across(everything(), as.numeric))

# Compute the correlation matrix
cor_matrix <- cor(df_, use = "complete.obs", method = "spearman")

# Plot the correlation matrix
ggcorrplot::ggcorrplot(
  cor_matrix,
  method = "circle",
  lab = TRUE,
  lab_size = 2,
  colors = c("red", "white", "green4"),
  title = "Correlation Matrix",
  hc.order = TRUE,
) +
  theme(
    axis.text.x = element_text(size = 8), # Adjust x-axis text size
    axis.text.y = element_text(size = 8) # Adjust y-axis text size
  )
```


------------------------------------------------------------------------

# Univariable statistics

## Setup

Define function.

```{r}
fit_model <- function(
    df = df_uni, cols = cols, is_sandwich = T) {
  # Initialize
  out <- list()

  for (i in seq_along(cols)) {
    # Create formula
    model <- as.formula(paste(CHOSEN_OUTCOME, "~", cols[i]))

    fit <- suppressWarnings(glm(
      model,
      data = df,
      family = binomial()
    ))

    if (is_sandwich) {
      # Calculate robust standard errors (sandwich)
      robust_se <- sandwich::vcovHC(fit, type = "HC0")
      # Compute robust confidence intervals
      fit_results <- lmtest::coeftest(fit, vcov. = robust_se)
    } else {
      fit_results <- fit
    }

    # Summarize model coefficients
    fit_summary <-
      fit_results %>%
      # broom does not support exponentiation after coeftest so do it manually
      broom::tidy(conf.int = T) %>%
      mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
      arrange(term == "(Intercept)", p.value) %>%
      rename(odds_ratio = estimate, z_value = statistic)

    # Stylize and print
    stylized <-
      fit_summary %>%
      rename(
        "Predictors" = term,
        "Odds Ratios (OR)" = odds_ratio,
        "SE" = std.error,
        "Z-scores" = z_value,
        "P-values" = p.value,
        "CI (low)" = conf.low,
        "CI (high)" = conf.high,
      ) %>%
      mutate(
        `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
        `SE` = round(`SE`, 2),
        `CI (low)` = round(`CI (low)`, 2),
        `CI (high)` = round(`CI (high)`, 2),
        `P-values` = round(`P-values`, 3),
      )

    out[[i]] <- stylized
  }
  return(out)
}
```


## Unadjusted - Original variables

- Fit logistic with no adjustments
- Use original feature definitions

```{r}
# Define cols
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Fit model
out <- fit_model(df = df_uni, cols = cols, is_sandwich = T)

# Create table
univariable_unadjusted <-
  out %>%
  bind_rows() %>%
  filter(Predictors != "(Intercept)") %>%
  filter(`CI (high)` < 50) %>%
  arrange(`P-values`)

# Print
univariable_unadjusted %>%
  sable()
```


## Adjusted - Original variables

- Fit logistic
- Adjust for age at first treatment and sex
- Use original feature definitions

```{r}
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Initialize
out <- list()
df <- df_uni

for (i in seq_along(cols)) {
  # Create formula
  model <- as.formula(paste(
    CHOSEN_OUTCOME,
    "~",
    "age_at_first_treatment_yrs + is_male +",
    cols[i]
  ))

  fit <- suppressWarnings(glm(
    model,
    data = df,
    family = binomial()
  ))

  # Calculate robust standard errors (sandwich)
  robust_se <- sandwich::vcovHC(fit, type = "HC0")
  # Compute robust confidence intervals
  robust_ci <- lmtest::coeftest(fit, vcov. = robust_se)

  # Summarize model coefficients
  fit_summary <-
    robust_ci %>%
    # broom does not support exponentiation after coeftest so do it manually
    broom::tidy(conf.int = T) %>%
    mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
    arrange(term == "(Intercept)", p.value) %>%
    rename(odds_ratio = estimate, z_value = statistic)

  # Stylize and print
  stylized <-
    fit_summary %>%
    rename(
      "Predictors" = term,
      "Odds Ratios (OR)" = odds_ratio,
      "SE" = std.error,
      "Z-scores" = z_value,
      "P-values" = p.value,
      "CI (low)" = conf.low,
      "CI (high)" = conf.high,
    ) %>%
    mutate(
      `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
      `SE` = round(`SE`, 2),
      `CI (low)` = round(`CI (low)`, 2),
      `CI (high)` = round(`CI (high)`, 2),
      `P-values` = round(`P-values`, 3),
    )

  out[[i]] <- stylized
}
```


Print results.

```{r}
# Create table
univariable_adjusted <-
  out %>%
  bind_rows() %>%
  filter(Predictors != "(Intercept)") %>%
  filter(!str_starts(Predictors, "age_at_first_treatment_yrs|is_male")) %>%
  filter(`CI (high)` < 50) %>%
  arrange(`P-values`)

# Print
univariable_adjusted %>%
  sable()
```


## Unadjusted - Recoded variables

- Fit logistic with no adjustments
- Use the recoded feature definitions (less levels for max statistical power)

```{r}
# Define columns
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Initialize
out <- list()
df <- df_multi

for (i in seq_along(cols)) {
  # Create formula
  model <- as.formula(paste(CHOSEN_OUTCOME, "~", cols[i]))

  fit <- suppressWarnings(glm(
    model,
    data = df,
    family = binomial()
  ))

  # Calculate robust standard errors (sandwich)
  robust_se <- sandwich::vcovHC(fit, type = "HC0")
  # Compute robust confidence intervals
  robust_ci <- lmtest::coeftest(fit, vcov. = robust_se)

  # Summarize model coefficients
  fit_summary <-
    robust_ci %>%
    # broom does not support exponentiation after coeftest so do it manually
    broom::tidy(conf.int = T) %>%
    mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
    arrange(term == "(Intercept)", p.value) %>%
    rename(odds_ratio = estimate, z_value = statistic)

  # Stylize and print
  stylized <-
    fit_summary %>%
    rename(
      "Predictors" = term,
      "Odds Ratios (OR)" = odds_ratio,
      "SE" = std.error,
      "Z-scores" = z_value,
      "P-values" = p.value,
      "CI (low)" = conf.low,
      "CI (high)" = conf.high,
    ) %>%
    mutate(
      `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
      `SE` = round(`SE`, 2),
      `CI (low)` = round(`CI (low)`, 2),
      `CI (high)` = round(`CI (high)`, 2),
      `P-values` = round(`P-values`, 3),
    )

  out[[i]] <- stylized
}
```


Print results.

```{r}
# Create table
univariable_unadjusted_recoded <-
  out %>%
  bind_rows() %>%
  filter(Predictors != "(Intercept)") %>%
  filter(`CI (high)` < 50) %>%
  arrange(`P-values`)

# Print
univariable_unadjusted_recoded %>%
  sable()
```


## Adjusted - Recoded

- Fit logistic with no adjustments
- Adjust for age at first treatment and sex
- Use the recoded feature definitions (less levels for max statistical power)

```{r}
# Define columns
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Initialize
out <- list()
df <- df_multi

for (i in seq_along(cols)) {
  # Create formula
  model <- as.formula(paste(
    CHOSEN_OUTCOME,
    "~",
    "age_at_first_treatment_yrs + is_male +",
    cols[i]
  ))

  fit <- suppressWarnings(glm(
    model,
    data = df,
    family = binomial()
  ))

  # Calculate robust standard errors (sandwich)
  robust_se <- sandwich::vcovHC(fit, type = "HC0")
  # Compute robust confidence intervals
  robust_ci <- lmtest::coeftest(fit, vcov. = robust_se)

  # Summarize model coefficients
  fit_summary <-
    robust_ci %>%
    # broom does not support exponentiation after coeftest so do it manually
    broom::tidy(conf.int = T) %>%
    mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
    arrange(term == "(Intercept)", p.value) %>%
    rename(odds_ratio = estimate, z_value = statistic)

  # Stylize and print
  stylized <-
    fit_summary %>%
    rename(
      "Predictors" = term,
      "Odds Ratios (OR)" = odds_ratio,
      "SE" = std.error,
      "Z-scores" = z_value,
      "P-values" = p.value,
      "CI (low)" = conf.low,
      "CI (high)" = conf.high,
    ) %>%
    mutate(
      `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
      `SE` = round(`SE`, 2),
      `CI (low)` = round(`CI (low)`, 2),
      `CI (high)` = round(`CI (high)`, 2),
      `P-values` = round(`P-values`, 3),
    )

  out[[i]] <- stylized
}
```


Print results.

```{r}
# Create table
univariable_adjusted_recoded <-
  out %>%
  bind_rows() %>%
  filter(Predictors != "(Intercept)") %>%
  filter(!str_starts(Predictors, "age_at_first_treatment_yrs|is_male")) %>%
  filter(`CI (high)` < 50) %>%
  arrange(`P-values`)

# Print
univariable_adjusted_recoded %>%
  sable()
```


------------------------------------------------------------------------

# Interaction analysis

## By hemorrhage

Fit logistic.

```{r}
# Define columns
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Remove unwanted columns
unwanted <- c("modified_rankin_score_final")
cols <- cols[!cols %in% unwanted]

adjustors <- c(
  "age_at_first_treatment_yrs",
  "is_male"
)

# Initialize
out <- list()
df <- df_multi
k <- 1

for (i in seq_along(cols)) {
  # Do not fit model for variables that we are adjusting for
  if (cols[i] %in% adjustors) {
    next
  }

  # Create formula
  model <- as.formula(paste(
    CHOSEN_OUTCOME,
    "~",
    "age_at_first_treatment_yrs + is_male +",
    cols[i],
    "* has_hemorrhage",
    sep = ""
  ))

  fit <- suppressWarnings(glm(
    model,
    data = df,
    family = binomial()
  ))

  # Calculate robust standard errors (sandwich)
  robust_se <- sandwich::vcovHC(fit, type = "HC0")
  # Compute robust confidence intervals
  robust_ci <- lmtest::coeftest(fit, vcov. = robust_se)

  # Summarize model coefficients
  fit_summary <-
    robust_ci %>%
    # broom does not support exponentiation after coeftest so do it manually
    broom::tidy(conf.int = T) %>%
    mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
    arrange(term == "(Intercept)", p.value) %>%
    rename(odds_ratio = estimate, z_value = statistic)

  # Stylize and print
  stylized <-
    fit_summary %>%
    rename(
      "Predictors" = term,
      "Odds Ratios (OR)" = odds_ratio,
      "SE" = std.error,
      "Z-scores" = z_value,
      "P-values" = p.value,
      "CI (low)" = conf.low,
      "CI (high)" = conf.high,
    ) %>%
    mutate(
      `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
      `SE` = round(`SE`, 2),
      `CI (low)` = round(`CI (low)`, 2),
      `CI (high)` = round(`CI (high)`, 2),
      `P-values` = round(`P-values`, 3),
    )

  out[[k]] <- stylized
  k <- k + 1
}
```


Print all results results.

```{r}
# Define values
unwanted <- c(
  "is_maleTRUE",
  "age_at_first_treatment_yrs",
  "(Intercept)",
  "has_hemorrhageTRUE"
)

# Initialize objects
isolated_values <- list()

# Get main and interaction effects
for (i in seq_along(out)) {
  # Get data
  result <- out[[i]]

  # Get all main effects of interest
  main_effects <-
    result %>%
    filter(!Predictors %in% unwanted) %>%
    filter(!str_detect(Predictors, ":")) %>%
    select(-"SE", -"Z-scores")

  # Get the interaction effects
  interaction_effects <-
    result %>%
    filter(!Predictors %in% unwanted) %>%
    filter(str_detect(Predictors, ":")) %>%
    mutate(Predictors = str_extract(Predictors, "^[^:]+")) %>%
    select(-"SE", -"Z-scores") %>%
    rename_with(~ paste("Interaction -", .), -Predictors)

  isolated_values[[i]] <-
    main_effects %>%
    left_join(interaction_effects, by = "Predictors")
}
```


Create and print table.

```{r}
# Create table
univariable_adjusted_recoded_interactions <-
  isolated_values %>%
  bind_rows() %>%
  arrange(`Interaction - P-values`)

# Print
univariable_adjusted_recoded_interactions %>%
  sable()
```

Eloquence.

```{r}
df %>%
  count(has_hemorrhage, is_eloquent_location, !!sym(CHOSEN_OUTCOME)) %>%
  drop_na() %>%
  arrange(has_hemorrhage, is_eloquent_location) %>%
  group_by(has_hemorrhage, is_eloquent_location) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```

Deficit.

```{r}
df %>%
  count(has_hemorrhage, has_deficit, !!sym(CHOSEN_OUTCOME)) %>%
  drop_na() %>%
  arrange(has_hemorrhage, has_deficit) %>%
  group_by(has_hemorrhage, has_deficit) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```

Seizures.

```{r}
df %>%
  count(has_hemorrhage, has_seizures, !!sym(CHOSEN_OUTCOME)) %>%
  drop_na() %>%
  arrange(has_hemorrhage, has_seizures) %>%
  group_by(has_hemorrhage, has_seizures) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```


## By SM grade high vs. low

Fit logistic.

```{r}
# Define columns
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL
))

# Remove unwanted columns
unwanted <- c("modified_rankin_score_final")
cols <- cols[!cols %in% unwanted]

adjustors <- c(
  "age_at_first_treatment_yrs",
  "is_male"
)

# Initialize
out <- list()
df <- df_multi
k <- 1

for (i in seq_along(cols)) {
  # Do not fit model for variables that we are adjusting for
  if (cols[i] %in% adjustors) {
    next
  }

  # Create formula
  model <- as.formula(paste(
    CHOSEN_OUTCOME,
    "~",
    "age_at_first_treatment_yrs + is_male +",
    cols[i],
    "* is_spetzler_martin_grade_less_than_4",
    sep = ""
  ))

  fit <- suppressWarnings(glm(
    model,
    data = df,
    family = binomial()
  ))

  # Calculate robust standard errors (sandwich)
  robust_se <- sandwich::vcovHC(fit, type = "HC0")
  # Compute robust confidence intervals
  robust_ci <- lmtest::coeftest(fit, vcov. = robust_se)

  # Summarize model coefficients
  fit_summary <-
    robust_ci %>%
    # broom does not support exponentiation after coeftest so do it manually
    broom::tidy(conf.int = T) %>%
    mutate(across(c(estimate, "conf.low", "conf.high"), exp)) %>%
    arrange(term == "(Intercept)", p.value) %>%
    rename(odds_ratio = estimate, z_value = statistic)

  # Stylize and print
  stylized <-
    fit_summary %>%
    rename(
      "Predictors" = term,
      "Odds Ratios (OR)" = odds_ratio,
      "SE" = std.error,
      "Z-scores" = z_value,
      "P-values" = p.value,
      "CI (low)" = conf.low,
      "CI (high)" = conf.high,
    ) %>%
    mutate(
      `Odds Ratios (OR)` = round(`Odds Ratios (OR)`, 2),
      `SE` = round(`SE`, 2),
      `CI (low)` = round(`CI (low)`, 2),
      `CI (high)` = round(`CI (high)`, 2),
      `P-values` = round(`P-values`, 3),
    )

  out[[k]] <- stylized
  k <- k + 1
}
```


Print all results results.

```{r}
# Define values
unwanted <- c(
  "is_maleTRUE",
  "age_at_first_treatment_yrs",
  "(Intercept)",
  "is_spetzler_martin_grade_less_than_4TRUE"
)

# Initialize objects
isolated_values <- list()

# Get main and interaction effects
for (i in seq_along(out)) {
  # Get data
  result <- out[[i]]

  # Get all main effects of interest
  main_effects <-
    result %>%
    filter(!Predictors %in% unwanted) %>%
    filter(!str_detect(Predictors, ":")) %>%
    select(-"SE", -"Z-scores")

  # Get the interaction effects
  interaction_effects <-
    result %>%
    filter(!Predictors %in% unwanted) %>%
    filter(str_detect(Predictors, ":")) %>%
    mutate(Predictors = str_extract(Predictors, "^[^:]+")) %>%
    select(-"SE", -"Z-scores") %>%
    rename_with(~ paste("Interaction -", .), -Predictors)

  isolated_values[[i]] <-
    main_effects %>%
    left_join(interaction_effects, by = "Predictors")
}
```


Create and print table.

```{r}
# Create table
univariable_adjusted_recoded_interactions <-
  isolated_values %>%
  bind_rows() %>%
  arrange(`Interaction - P-values`)

# Print
univariable_adjusted_recoded_interactions %>%
  sable()
```

Eloquence.

```{r}
df %>%
  count(
    is_spetzler_martin_grade_less_than_4,
    is_eloquent_location,
    !!sym(CHOSEN_OUTCOME)
  ) %>%
  drop_na() %>%
  arrange(is_spetzler_martin_grade_less_than_4, is_eloquent_location) %>%
  group_by(is_spetzler_martin_grade_less_than_4, is_eloquent_location) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```

Deficit.

```{r}
df %>%
  count(
    is_spetzler_martin_grade_less_than_4,
    has_deficit,
    !!sym(CHOSEN_OUTCOME)
  ) %>%
  drop_na() %>%
  arrange(is_spetzler_martin_grade_less_than_4, has_deficit) %>%
  group_by(is_spetzler_martin_grade_less_than_4, has_deficit) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```

Seizures.

```{r}
df %>%
  count(
    is_spetzler_martin_grade_less_than_4,
    has_seizures,
    !!sym(CHOSEN_OUTCOME)
  ) %>%
  drop_na() %>%
  arrange(is_spetzler_martin_grade_less_than_4, has_seizures) %>%
  group_by(is_spetzler_martin_grade_less_than_4, has_seizures) %>%
  mutate(
    num_total = sum(n),
    pct_total = scales::percent(n / num_total, 1)
  ) %>%
  sable()
```


------------------------------------------------------------------------

# Multivariable statistics

## Resources

Read the following guides:

- [How to use glmnet](https://glmnet.stanford.edu/articles/glmnet.html)

- [Selective inference using forward selection](https://stephens999.github.io/misc/selective_inference_toy.html)

- [LASSO when conditioning on less: ROSI](https://selective-inference.github.io/Python-software/algorithms/ROSI.html)


## Data prep

Define unwanted columns.

```{r}
# Spetzler-Martin (SM) grade = size + eloquence + deep venous drainage.
# Two encodings of the grade itself: the granular 1-5 and the dichotomized SM<4.
SM_GRADES <- c(
  "spetzler_martin_grade",
  "is_spetzler_martin_grade_less_than_4"
)

# The individual SM components (some dimensions are encoded twice)
SM_COMPONENTS <- c(
  # Size
  "size_score",
  "max_size_cm",
  # Eloquence
  "is_eloquent_location",
  "location",
  # Venous drainage (has_deep_venous_drainage is identical to venous_drainage)
  "venous_drainage",
  "has_deep_venous_drainage"
)

# Lawton-Young = the Supplemented SM grade: SM plus age, hemorrhagic
# presentation and diffuse nidus, pre-computed as one score. It ALREADY contains
# SM, so it is the most comprehensive grade.
LY_GRADE <- "lawton_young_grade"

# What Lawton-Young adds on top of SM
LY_EXTRAS <- c(
  "has_hemorrhage",
  "is_diffuse_nidus"
  # LY includes age, but it lumps all <20yo into one bucket - all our patients are <18yo
  # "age_at_first_treatment_yrs"
)

# When a component is encoded twice, keep the preferred version and drop the
# duplicate, to avoid collinearity between two encodings of the same thing.
REDUNDANT_COMPONENTS <- c(
  # Size: keep the granular max_size_cm instead
  "size_score",
  # Eloquence: keep is_eloquent_location instead (more relevant)
  "location",
  # Venous drainage: keep has_deep_venous_drainage instead (more explainable name)
  "venous_drainage"
)

# Columns excluded from every model
unwanted_all <- c(
  # These mRS are recorded AFTER tx, so using them would leak the outcome
  "modified_rankin_score_postop_within_1_week",
  "modified_rankin_score_final",
  # Keep the pre-tx mRS (the baseline closest to treatment); this earlier is redundant
  "modified_rankin_score_presentation",
  # Treatment-era flags: proxies for calendar time, not the clinical question
  "is_first_tx_after_2009",
  "is_first_tx_from_2001"
)

# If the outcome itself is grade-related, drop the grades so they cannot leak it
if (str_detect(params$TITLE, "grade")) {
  unwanted_all <- c(unwanted_all, SM_GRADES, LY_GRADE)
}

# If the outcome is hemorrhage-related, drop hemorrhage-at-presentation
if (str_detect(params$TITLE, "Hemorrhage")) {
  unwanted_all <- c(unwanted_all, "has_hemorrhage")
}

# all: nothing extra removed, so it keeps both SM forms, Lawton-Young and every
# component. Only the collinearity-immune LASSO methods use this set.

# Components model: keep the granular components, drop every composite grade
unwanted_without_gradings <- unique(c(
  unwanted_all,
  SM_GRADES,
  LY_GRADE,
  REDUNDANT_COMPONENTS
))

# Grades model: keep the Supplemented SM grade (Lawton-Young) alone. Drop both
# SM forms (redundant with LY) and every component.
unwanted_with_gradings <- unique(c(
  unwanted_all,
  SM_COMPONENTS,
  SM_GRADES,
  LY_EXTRAS
))
```


Create dataset.

```{r}
# Define cols
cols <- unname(c(
  EXPOSURES_CONTINUOUS,
  EXPOSURES_BINARY,
  EXPOSURES_CATEGORICAL,
  CHOSEN_OUTCOME
))

# Define column sets for each analysis
cols_all <- cols[!cols %in% unwanted_all]
cols_with_gradings <- cols[!cols %in% unwanted_with_gradings]
cols_without_gradings <- cols[!cols %in% unwanted_without_gradings]

# Create df of interest
df_all <-
  # Use the non-recoded dataset
  df_multi %>%
  dplyr::select(all_of(cols_all)) %>%
  drop_na()

df_with_gradings <-
  # Use the non-recoded dataset
  df_multi %>%
  dplyr::select(all_of(cols_with_gradings)) %>%
  drop_na()

df_no_scores <-
  # Use the non-recoded dataset
  df_multi %>%
  dplyr::select(all_of(cols_without_gradings)) %>%
  drop_na()

# Create formula
frla <- as.formula(paste(CHOSEN_OUTCOME, " ~ . - 1"))

# Create matrices with all variables
# (remove falsely created FALSE columns)
X_ <- model.matrix(frla, df_all)
X_all <- X_[, !grepl("FALSE$", colnames(X_))]
y_all <- df_all %>%
  pull(CHOSEN_OUTCOME) %>%
  as.numeric()

# Create matrices with scores
X_ <- model.matrix(frla, df_with_gradings)
X_with_gradings <- X_[, !grepl("FALSE$", colnames(X_))]
y_with_gradings <- df_with_gradings %>%
  pull(CHOSEN_OUTCOME) %>%
  as.numeric()

# Create matrices with score components
X_ <- model.matrix(frla, df_no_scores)
X_without_gradings <- X_[, !grepl("FALSE$", colnames(X_))]
y_without_gradings <- df_no_scores %>%
  pull(CHOSEN_OUTCOME) %>%
  as.numeric()


# Remove duplicate columns
# Define priority pairs (keep first, remove second)
priority_pairs <- list(
  c("is_maleTRUE", "procedure_combinationsS")
  # Add more pairs as needed
)

# Remove lower priority columns when both exist
for (pair in priority_pairs) {
  if(all(pair %in% colnames(X_all))) {
    cols_ <- colnames(X_all)
    X_all <- X_all[, !cols_ %in% pair[2]]
  }
  if(all(pair %in% colnames(X_with_gradings))) {
    cols_ <- colnames(X_with_gradings)
    X_with_gradings <- X_with_gradings[, !cols_ %in% pair[2]]
  }
  if(all(pair %in% colnames(X_without_gradings))) {
    cols_ <- colnames(X_without_gradings)
    X_without_gradings <- X_without_gradings[, !cols_ %in% pair[2]]
  }
}


# X should be centered for LASSO (necessary for glmnet)
# (scale = FALSE as this is done by the standardization step in glmnet)
center_values <- function(X) {
  # Identify binary columns (those with only 0s and 1s)
  binary_cols <- apply(X, 2, function(x) all(x %in% c(0, 1)))
  # Center but do not scale - avoid centering one-hot encoded vectors
  X[, !binary_cols] <- scale(X[, !binary_cols], center = T, scale = F)
  # Return
  return(X)
}

X_all_centered <- center_values(X_all)
X_with_gradings_centered <- center_values(X_with_gradings)
X_without_gradings_centered <- center_values(X_without_gradings)

# See the names of X to match column numbers to column names later on
colnames(X_without_gradings_centered)
```


## Configs

```{r}
# Drop the surgery dummy when present (collinear with the outcome)
drop_surgery <- function(x) x[, !colnames(x) %in% "is_surgeryTRUE", drop = FALSE]

# Which methods to run (all TRUE by default)
METHODS <- list(
  # Selective-inference LASSO at the optimal lambda and the lambda at min loss
  selective_lasso = TRUE,
  # Bootstrap of the selective-inference LASSO
  bootstrap = TRUE,
  # Multi-sample splitting with cross-fitting
  split = TRUE,
  # Firth penalized logistic on the selected set
  firth = TRUE,
  # Forward step-wise (logistic ORs + native linear inference)
  stepwise = TRUE,
  # Univariable logistic, one predictor at a time
  univariable = TRUE,
  # p-value-screened multivariable logistic
  multivariable = TRUE
)

CONDITIONS <- list(
  # All predictors (collinear → no multivariable)
  all = list(
    x = X_all_centered,
    y = y_all,
    multi_unwanted = NULL
  ),
  # Score components without the composite scores
  without_gradings = list(
    x = X_without_gradings_centered,
    y = y_without_gradings,
    multi_unwanted = unwanted_without_gradings
  ),
  # Composite scores without their components
  with_gradings = list(
    x = X_with_gradings_centered,
    y = y_with_gradings,
    multi_unwanted = unwanted_with_gradings
  ),
  # As without_gradings, drop is_surgery (no multivariable)
  without_gradings_no_surgery = list(
    x = drop_surgery(X_without_gradings_centered),
    y = y_without_gradings,
    multi_unwanted = NULL
  ),
  # As with_gradings, drop is_surgery (no multivariable)
  with_gradings_no_surgery = list(
    x = drop_surgery(X_with_gradings_centered),
    y = y_with_gradings,
    multi_unwanted = NULL
  )
)

# Traditional multivariable regression additional parameters
MULTI_PARAMS <- list(
  # Data frame for the multivariable fit (non-recoded)
  df = df_multi,
  # Outcome column name
  outcome = CHOSEN_OUTCOME,
  # Univariable tables the predictor screen binds and filters
  screen_tables = list(
    univariable_adjusted_recoded,
    univariable_unadjusted_recoded
  ),
  # Univariable p-value threshold to enter the screen
  pval_threshold = 0.2,
  # Columns the multivariable fit requires complete
  drop_na_cols = c("modified_rankin_score_pretreatment", "location")
)

# Univariable table to report (unadjusted, recoded)
UNI_TABLE <- univariable_unadjusted_recoded

# Which methods to render forest plots for (all TRUE by default)
PLOT_METHODS <- list(
  # Selective LASSO at the optimal (stable-plateau) lambda
  lasso_opt = TRUE,
  # Selective LASSO at lambda.min
  lasso_min = TRUE,
  # Multi-sample splitting with cross-fitting
  split = TRUE
)
```


## Parallelize

```{r}
# furrr/future power the bootstrap and split loops. Leaves one core free for the OS
if (PARALLELIZE) {
  future::plan(future::multisession, workers = future::availableCores() - 1)
}
```


## Methods

```{r}
source("multivariable-methods.R")
```


## Run

```{r}
# Compute all multivariable regression methods at once
results <- compute_all_methods(
  # Each condition's centered LASSO design + its multivariable exclusion set
  conditions = CONDITIONS,
  # Pre-computed univariable table (attached to the "all" rows)
  uni_table = UNI_TABLE,
  # Bundled multivariable settings (computed per condition inside)
  multi_params = MULTI_PARAMS,
  # Method on/off toggles (missing entries default to TRUE)
  methods = METHODS,
  # Selective-inference significance level (95% CI at 0.05, plus 90% companion)
  si_alpha = 0.05,
  # Master RNG seed; reset at the start of every stochastic method and condition
  seed = 141845,
  # Cross-validation folds for the LASSO
  nfolds = 10,
  # Random half-splits for the split-and-cross-fit method
  n_splits = 1000,
  # Bootstrap replicates
  boot_B = 1000,
  # Bootstrap CI significance level
  boot_alpha = 0.05,
  # Bootstrap lambda rule (0 = lambda.min, 1 = lambda.1se)
  boot_k = 0,
  # Minimum selection frequency (%) for a variable to appear in the split plots
  split_min_pct = 5
)

# Results table
df_combined <- results$combined
```


Release the parallel worker sessions.

```{r}
if (PARALLELIZE) {
  future::plan(future::sequential)
}
```


Inspect results.

```{r}
df_combined %>% sable()
```


## Plot

```{r}
# Build a forest plot for each enabled selective-LASSO/split model
plots <- plot_all_methods(
  # The computed results (uses its plot_tables element)
  results = results,
  # Named lookup that maps raw term names to readable labels
  col_names = c(EXPOSURES_CONTINUOUS, EXPOSURES_BINARY, EXPOSURES_CATEGORICAL),
  # Which methods to plot (missing entries default to TRUE)
  methods = PLOT_METHODS
)
```


------------------------------------------------------------------------

# Write

## Setup

Create the necessary directories.

```{r, results="hold"}
# Get today's date
today <- Sys.Date()
today <- format(today, "%Y-%m-%d")

# Create folder names
base_folder <- file.path(DST_DIRNAME, ANALYSIS_NAME)
date_folder <- file.path(base_folder, today)
csv_folder <- file.path(date_folder, "csv")
pdf_folder <- file.path(date_folder, "pdf")
html_folder <- file.path(date_folder, "html")
models_folder <- file.path(date_folder, "rds")

# Create folders
suppressWarnings(dir.create(base_folder))
suppressWarnings(dir.create(date_folder))
suppressWarnings(dir.create(csv_folder))
suppressWarnings(dir.create(pdf_folder))
suppressWarnings(dir.create(html_folder))
suppressWarnings(dir.create(models_folder))

# Print folder names
print(base_folder)
print(date_folder)
print(csv_folder)
print(pdf_folder)
print(html_folder)
```


## Models

### Bootstrap

```{r}
# Save the bootstrap draws so the resampling is reproducible later
if (!is.null(results$raw_all)) {
  saveRDS(results$raw_all, file.path(models_folder, "bootstrap_raw_all.rds"))
}
```


## Figures

### Visualizations

```{r}
# Save
ggsave(
  file.path(pdf_folder, "sm_grade_vs_complications_bar_chart.pdf"),
  plot = plots$sm_grade_vs_complications_bar_chart,
  width = 8,
  height = 6
)
```


### Multivariable statistics

```{r}
# Export every forest to forest_<label>_<condition>.pdf
write_all_plots(plots, plot_folder = pdf_folder)
```


## Tables

### Univariable statistics

```{r}
write_csv(
  univariable_unadjusted,
  file.path(csv_folder, "univariable_unadjusted.csv")
)

write_csv(
  univariable_adjusted,
  file.path(csv_folder, "univariable_adjusted.csv")
)

write_csv(
  univariable_adjusted_recoded,
  file.path(csv_folder, "univariable_adjusted_recoded.csv")
)

write_csv(
  univariable_adjusted_recoded_interactions,
  file.path(csv_folder, "univariable_adjusted_recoded_interactions.csv")
)
```


### Multivariable statistics

```{r}
# Write every computed table to its CSV (export only; no computation here)
write_all_methods(results, csv_folder)
```


------------------------------------------------------------------------


# Reproducibility

## Linting and styling

```{r, eval=FALSE}
# Style current file
styler::style_file(
  path = rstudioapi::getSourceEditorContext()$path,
  style = styler::tidyverse_style
)

# Lint current file
lintr::lint(rstudioapi::getSourceEditorContext()$path)
```


## Dependency management

```{r, eval=FALSE, results="hide"}
# Clean up project of libraries not in use
# (use prompt = FALSE to avoid the interactive session)
# (packages can only be removed in interactive mode b/c this is destructive)
renv::clean(prompt = TRUE)

# Update lock file with new packages
renv::snapshot()
```


## Containerization

```{r}
# Only run this if option is set to TRUE
if (UPDATE_DOCKERFILE) {
  # Create a dockerfile from the session info
  my_dockerfile <- containerit::dockerfile(from = sessionInfo(), env = ls())
  # Write file
  write(my_dockerfile, file = "~/Dockerfile")
  # Print
  print(my_dockerfile)
}
```


------------------------------------------------------------------------

# Documentation {.tabset}

## Session Info

```{r session_info, echo=FALSE}
print(sessionInfo(), locale = FALSE)
```

## References

```{r refs, echo=FALSE}
(.packages()) %>%
  sort() %>%
  lapply(citation) %>%
  lapply(c) %>%
  unique()
```
